---
title: "Homework 01"
author: "Jinhai Chen,3170104884"
date: "06/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Homework for July 6th

### Solution for Exercise 1

a.
```{r}
  iowa.df <- read.csv("Data/iowa.csv",header = T,sep=";")
```

b.
```{r}
  dim(iowa.df)
```

c.
```{r}
  colnames(iowa.df)
```

d.
```{r}
  iowa.df[5,7]
```

e.
```{r}
  iowa.df[2,]
```

### Solution for Exercise 2

a.
```{r}
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
#sum(vector1)
```
The first computation can lead out to results, while the last two can not. The reason is that vector1 consists of 4 character variable, but the sort and sum can only sort or add up the numbers rather than the characters. The following codes will do.
```{r}
vec1 <-c(5,12,7,32)
max(vec1)
sort(vec1)
sum(vec1)
```

b.

```{r}
vector2 <- c("5",7,12)
#vector2[2] + vector2[3]

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
#list4[2]+list4[4]
```
The first question is that the type of vector2 is
```{r}
typeof(vector2)
```
as a resulr of which can not be added up. the summation can only work for numbers as follows:

```{r}
  vec2 <- c(5,7,12)
  vec2[2] + vec2[3]
```