---
title: "Homework 01"
author: "Jinhai Chen,3170104884"
date: "06/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Homework for July 6th

### Solution for Exercise 1

a.
```{r}
  iowa.df <- read.csv("Data/iowa.csv",header = T,sep=";")
```

b.
```{r}
  dim(iowa.df)
```

c.
```{r}
  colnames(iowa.df)
```

d.
```{r}
  iowa.df[5,7]
```

e.
```{r}
  iowa.df[2,]
```

### Solution for Exercise 2

a.
```{r}
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
#sum(vector1)
```
The first computation can lead out to results, while the last two can not. The reason is that vector1 consists of 4 character variable, but the sort and sum can only sort or add up the numbers rather than the characters. The following codes will do.
```{r}
vec1 <-c(5,12,7,32)
max(vec1)
sort(vec1)
sum(vec1)
```

b.

```{r}
vector2 <- c("5",7,12)
#vector2[2] + vector2[3]

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
#list4[2]+list4[4]
```
The first question is that the type of vector2 is
```{r}
typeof(vector2)
```
as a resulr of which can not be added up. the summation can only work for numbers as follows:
```{r}
  vec2 <- c(5,7,12)
  vec2[2] + vec2[3]
```
Second, function data.frame creates coupled collections of variables like lists or matrices, which in that way makes the summation available.
Last, [] and [[]] are used to access elements in a list: [] gain access the name and the value, while [[]] gain access to only the number. So list4[[2]]+list4[[4]] returns a nummber of 168 while the other returns an error.

### Solution for Exercise 3
a.
```{r}
  seq(1,10000,by=372)
  seq(1,10000,length.out=50)
```

b.
```{r}
rep(1:3, times=3)
rep(1:3, each=3)
```
The first one repeats the sequence from 1 to 3 for three times while the second one repeats each elements for three times inside the sequence.

### Solution for MB.Ch1.2
```{r}
  library(DAAG)
  orings <- DAAG::orings
  dataframe1 <- orings[c(1,2,4,11,13),]
  plot(dataframe1[,1],dataframe1[,4],xlab = "Temperature",ylab = "Total incidents")
  plot(orings[,1],orings[,4],xlab = "Temperature",ylab = "Total incidents")
```

### Solution for MB.Ch1.4
a.
```{r}
ais <- DAAG::ais
str(ais)
sum(is.na(ais))
```

b.
```{r}
table <- table(ais$sex,ais$sport)
table
ratio = table[1,]/table[2,]
names(ratio[(ratio>2) | (ratio<1/2)])
```

### Solution for MB.Ch1.6
```{r}
elevation <- c(217, 254,248,254,253,227,178,207,217)
area <- c(24387,5374,4624,2247,1353,1223,1151,755,657)
names <- c("Winnipeg","Winnipegosis","Manitoba","SouthernIndian","Cedar","Island","Gods","Corss","Playgreen")
Manitoba.lakes <- data.frame("elevatoin"=elevation,"area"=area)
row.names(Manitoba.lakes) <- names
```

a.
```{r}
attach(Manitoba.lakes)
plot(log2(area) ~ elevation, pch=16, xlim=c(170,280))
# NB: Doubling the area increases log2(area) by 1.0
text(log2(area) ~ elevation, labels=row.names(Manitoba.lakes), pos=4)
text(log2(area) ~ elevation, labels=area, pos=2) 
title("Manitoba’s Largest Lakes")
```

b.
```{r}
plot(area ~ elevation, pch=16, xlim=c(170,280), ylog=T)
text(area ~ elevation, labels=row.names(Manitoba.lakes), pos=4, ylog=T)
text(area ~ elevation, labels=area, pos=2, ylog=T) 
title("Manitoba’s Largest Lakes")

```

### Solution for MB.Ch1.7
a.
```{r}
dotchart(area,xlab="area",ylab="lakes",labels="names")
title("The areas of the Manitoba lakes on a linear scale")
```

b.
```{r}
dotchart(log2(area),xlab="log2(area)",ylab="lakes",labels="names")
title("The areas of the Manitoba lakes on logarithmic scale")
```

### Solution for MB.Ch1.8
```{r}
sum(area)
```